Computing in the Field of Rationals

### 2012-06-24 rational_ext.tex
_

@@

  {{c v |--> v'}} iff {{<c, v, []> |-->* [c, v', [] ]}}

@@

\end{enumerate}
\end{proposition}
\begin{proof}
  For convenience let us draw {{swap* (;) eps*}} shown on the left
  instead as shown on the right.

@@

The two coherence conditions are more intuitive when presented
graphically. They require that the combinators below be equal to
{{id}}, at the appropriate type (see Sec. 4
\cite{springerlink:10.1007/978-3-642-12821-94}).

@@

%%%%%%%%%%%%%%%%%%%
\subsection{Booleans and Conditionals}

Given any combinator {{c : b <-> b}} we can construct a combinator called
{{if_c : bool*b <->bool*b}} in terms of {{c}}, where {{if_c}} behaves like a
one-armed $\mathit{if}$-expression. If the supplied boolean is {{true}} then
the combinator {{c}} is used to transform the value of type~{{b}}. If the
boolean is {{false}}, then the value of type {{b}} remains unchanged. We can
write down the combinator for {{if_c}} in terms of {{c}} as 
{{ dist (;) ((id (*) c) (+) id) (;) factor }}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cloning}

Although cloning is generally not allowed in reversible languages, it is
possible at the cost of having additional constant inputs. For example,
consider the gate {{else_{not} }}. Generally, the gate maps
{{(false,a)}} to {{(false,not a)}} and {{(true,a)}} to {{(true,a)}}. Focusing
on the cases in which the second input is {{true}}, we get that the gate maps
{{(false,true)}} to {{(false,false)}} and {{(true,true)}} to {{(true,true)}},
i.e., the gate clones the first input. A circuit of {{n}} parallel
{{else_{not} }} gates can hence clone {{n}} bits.  They also consume {{n}}
{{true}} inputs in the process.  Let us call this construction
{{clone^n_{bool} }}.

>>>>

%%%%%%%%%%%%%%%%%%%
\subsection{Constructions}

\paragraph*{Booleans and Conditionals.} 
Given any combinator {{c : b <-> b}} we can construct a combinator called
{{if_c : bool*b <->bool*b}} in terms of {{c}}, where {{if_c}} behaves like a
one-armed $\mathit{if}$-expression. If the supplied boolean is {{true}} then
the combinator {{c}} is used to transform the value of type~{{b}}. If the
boolean is {{false}}, then the value of type {{b}} remains unchanged. We can
write down the combinator for {{if_c}} in terms of {{c}} as 
{{ dist (;) ((id (*) c) (+) id) (;) factor }}.

\paragraph*{Cloning.}
Although cloning is generally not allowed in reversible languages, it is
possible at the cost of having additional constant inputs. For example,
consider the gate {{else_{not} }}. Generally, the gate maps
{{(false,a)}} to {{(false,not a)}} and {{(true,a)}} to {{(true,a)}}. Focusing
on the cases in which the second input is {{true}}, we get that the gate maps
{{(false,true)}} to {{(false,false)}} and {{(true,true)}} to {{(true,true)}},
i.e., the gate clones the first input. A circuit of {{n}} parallel
{{else_{not} }} gates can hence clone {{n}} bits.  They also consume {{n}}
{{true}} inputs in the process.  Let us call this construction
{{clone^n_{bool} }}.

@@

  {{c v |--> v'}} iff {{<c, v, []> |-->* [c', v', [] ]}}
\end{proposition}
\begin{proof}
  To prove the above, we first show that a more general lemma holds,
  namely that: {{c v |--> v'}} iff {{<c, v, C> |-->* [c', v', C ]}}.

  To show the left-to-right direction we proceed by induction on the
  derivation of {{c v |--> v'}}. In the case of primitive isomorphisms
  the condition holds trivially. In the case of composition, we work
  out the case of {{c1+c2}} as an example. Given {{c1 +c2 v |--> v'}}
  we have to show that there is a small step derivation sequence that
  matches it. Here {{v:b1+b2}} can be of the form {{left v1}} or
  {{right v2}}. Assuming {{left v1}}, we have:

%subcode{proof} include main
%@ c1 v1 |--> v1' ==> <c1, v1, C'> |-->* {[c1',v1',C']}
%@@ c1+c2 (left v1) |--> left v1' ==> ?

Choosing {{C' =Lft c2 C}}, we have the required derivation sequence: 
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1',v1', LeftP c2 C]} |--> {[c1'+c2,left v1',C]} }}. The proof follows similarly for the {{right v2}} case.

To show the right-to-left direction we proceed by induction of the
sequence of {{|-->*}} derivations. Again the case for primtive
isomorphisms follows trivially. Taking the case of {{c1+c2}} we are
required to show that given a sequence 
{{<c1+c2,v,C> |-->* {[c', v', C]} }} there is a derviation tree for
{{c1+c2 v |--> v'}}. In the case that {{v:v1+b2}} has the form
{{left~v1}}, this follows by observing that the given sequence must have the form
{{ <c1+c2, left v1, C> |--> <c1,v1,LeftP c2 C> |-->* {[c1',v1', LeftP c2 C]} |--> {[c1'+c2,left v1',C]} }}. 
For the strictly smaller inner subsequence by induction we have
{{c1~v1 |--> v1'}} which lets us complete the derivation of
{{c1+c2~(left v1) |--> left v1'}}. The {{right v2}} follows similarly.

@@

\end{enumerate}
\end{proposition}
\begin{proof}
For convenience, we draw the combinator {{swap* (;) eps*}} (shown on
the left) as shown on the right.

@@

\noindent
The coherence conditions are more intuitive when presented
graphically. They require that the combinators below be equal to
{{id}}, at the appropriate type (see Sec. 4
\cite{springerlink:10.1007/978-3-642-12821-94}).

@@
